Preprocessing and exploration of the narwhal data for the analysis with smoothSDE

### 2024-05-08 R/preprocessing/data_preprocessing.Rmd
preproseccing results with modified land polygons

@@

#get the coastline geometry from the geojson file
coastline_utm<-st_read(file.path(greenland_data_path,"coastline_utm.geojson"))
st_crs(coastline_utm)="+init=EPSG:32626"

@@

@@

```{r plot coastline and land}
plot_land=ggplot()+geom_sf(data=land_utm$geometry,fill=NA,col="brown")+
  coord_sf(datum=st_crs(32626))
plot_coastline=ggplot()+geom_sf(data=coastline_utm$geometry,fill=NA,col="blue")+
  coord_sf(datum=st_crs(32626))
```

@@

```{r display coastline , fig.cap="Coastline polygons and linestrings"}
plot_coastline
```

@@

res_land=nearest_points(positions=narwhal_positions,coastline=land_utm)
res_coastline=nearest_points(positions=narwhal_positions,coastline=coastline_utm)
})

@@

```{r compare distances to shore}
plot(res_land$min_distances-dataprep$DistanceShore)
plot(res_coastline$min_distances-dataprep$DistanceShore)
```

```{r summary distances to shore}
summary(res_land$min_distances)
summary(res_coastline$min_distances)
summary(dataprep$DistanceShore)
```
Coastline distances are a bit closer to original distances in the dataset.
Anyway, we shall keep distance computed from the land in the sequel, since the land polygons will be the reference domain
for constrained movement of the narwhals.

@@

df=data.frame(DistanceShore=res_coastline$min_distances[res_coastline$min_distances>0])
histo2=ggplot(df, aes(x =DistanceShore)) +
  geom_histogram(fill="cyan",bins=100,color="black")+xlab("Computed distance from coastline (km)")

df=data.frame(DistanceShore=res_land$min_distances[res_land$min_distances>0])
histo3=ggplot(df, aes(x =DistanceShore)) +
  geom_histogram(fill="cyan",bins=100,color="black")+xlab("Computed distance fom land (km)")
```

```{r display histograms}
histo1
histo2
histo3
```

```{=tex}
There are pretty big differences between distances calculates with the coastline data and the land data.

For the rest of the analysis, we'll keep the bounded domain defined by the polygons in \texttt{land\_utm} sf object as a reference.
Doing that involves having some points on land, that is points with Distance to shore equal to 0.
This should be essentially due to the error in the land polygons than to the error in GPS measurements.
```

@@

cat("Percentage of points on land with distance from coastline : \n",
    length(which(res_coastline$min_distances==0))/length(dataprep$time)*100)
```

@@

```{r check original points on land}
check_plot1=ggplot()+geom_sf(data=land_utm$geometry,fill=NA)+
    coord_sf(datum=st_crs(32626))+
  geom_point(data=dataprep,aes(x,y,col=(DistanceShore==0)),size=0.2)+
    scale_color_viridis_d()

@@

check_plot2=ggplot()+geom_sf(data=land_utm$geometry,fill=NA)+
    coord_sf(datum=st_crs(32626))+
  geom_point(data=df,aes(x,y,col=(DistanceShore==0)),size=0.2)+
    scale_color_viridis_d()

@@

df=data.frame(DistanceShore=res_land$min_distances[res_land$min_distances<2 & res_land$min_distances>0  ])
histo=ggplot(df, aes(x =DistanceShore)) +
  geom_histogram(fill="cyan",bins=100,color="black")+xlab("Computed distance to shore (km)")

@@

```{=tex}
$9 \%$ of the points are on land.
If the points are too close to then land (let's say distance to shore < 50m), the covariates AngleShore and Angle Normal are likely to be irrelevant.
```

>>>>

```{r save points as geojson in data folder}
points_sf <- st_as_sf(dataprep[,c("x","y")], coords = c("x","y"))
st_crs(points_sf)="+init=EPSG:32626 +units=km"
st_write(points_sf,file.path(narwhal_data_path,"narwhals_positions.shp"))
```

@@

@@

#get land polygons
modified_land_utm<-st_read(file.path(greenland_data_path,"modified_land_utm.shp"))
modified_land_utm <- st_transform(modified_land_utm, crs = "+init=EPSG:32626 +units=km")

#scoresby sound google
land_google_utm<-st_read(file.path(greenland_data_path,"land_google_utm.shp"))
st_crs(land_google_utm)="+init=EPSG:32626"
```

@@

```{r plot coastline and land}
plot_land=ggplot()+geom_sf(data=land_utm$geometry,fill="grey",col="brown")+
  coord_sf(datum=st_crs(32626))
plot_modified_land=ggplot()+geom_sf(data=modified_land_utm$geometry,fill="grey",col="blue")+
  coord_sf(datum=st_crs(32626))
plot_land_google=ggplot()+geom_sf(data=land_google_utm$geometry,fill="grey",col="blue")+
  coord_sf(datum=st_crs(32626))
```

@@

```{r display coastline , fig.cap="Coastline polygons and linestrings"}
ggplotly(plot_modified_land)
```

```{r}
ggplotly(plot_land_google)
```

@@

res_land=nearest_points(positions=narwhal_positions,coastline=land_utm)
})

@@

```{r compare distances to shore}
plot(res_land$min_distances-dataprep$DistanceShore)
```

```{r summary distances to shore}
summary(res_land$min_distances)
summary(dataprep$DistanceShore)
```

@@

df=data.frame(DistanceShore=res_land$min_distances[res_land$min_distances>0])
histo2=ggplot(df, aes(x =DistanceShore)) +
  geom_histogram(fill="cyan",bins=100,color="black")+xlab("Computed distance fom land (km)")
```

```{r display histograms}
histo1
histo2
```

```{=tex}

For the rest of the analysis, we'll keep the bounded domain defined by the polygons in \texttt{modified\_land\_utm} sf object as a reference.
Doing that involves having some points on land, that is points with Distance to shore equal to 0.
This should be essentially due to the error in the land polygons than to the error in GPS measurements.
```

@@

cat("Percentage of points on land with distance from modified : \n",
    length(which(res_modified_land$min_distances==0))/length(dataprep$time)*100)
```

@@

```{r check original points on land}
check_plot1=ggplot()+geom_sf(data=land_google_utm$geometry,fill="grey")+
    coord_sf(datum=st_crs(32626))+
  geom_point(data=dataprep,aes(x,y,col=(DistanceShore<0.05)),size=0.2)+
    scale_color_viridis_d()

@@

check_plot2=ggplot()+geom_sf(data=land_utm$geometry,fill="grey")+
    coord_sf(datum=st_crs(32626))+
  geom_point(data=df,aes(x,y,col=(DistanceShore==0)),size=0.2)+
    scale_color_viridis_d()

@@

df=data.frame(DistanceShore=res_land$min_distances[res_land$min_distances<0.2 & res_land$min_distances>0  ])
histo=ggplot(df, aes(x =DistanceShore)) +
  geom_histogram(fill="cyan",bins=100,color="black")+xlab("Computed distance to shore (km)")

@@
